Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-07-09 doc/main_jfm.tex
updates
updated Section 2

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\uu}{\mathbf{u}}
\renewcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}

@@

\author{
Szu-Pei Fu\aff{1},
Bryan Quaife\aff{2},
Rolf Ryham\aff{1}, \and
Yuan-Nan Young\aff{3}
}
 \affiliation{
\aff{1} Department of Mathematics, \\Fordham University, Bronx, New York 10458, USA
\aff{2}Department of Scientific Computing, \\Florida State University, Tallahassee, Florida 32306, USA
\aff{3}Department of Mathematical Sciences, New Jersey Institute of Technology,\\ Newark, New Jersey 07102, USA
 }

@@

\section{Governing  Equations}
\subsection{Hydrophobic Attraction Potential (HAP)}

The Hydrophobic attraction potential (HAP) is given by
\begin{equation}
\label{eq:main}
\Phi(\Omega,f) = \gamma  \min_{u \in \mathcal{A}}
 I[u]  = \int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2 \,dx,
\end{equation}
%
where $\gamma$ the interfacial tension, $\rho$ the screened length, and this domain functional describes the energy associated with hydrophobic surfaces~\cite{Fu20}. From the first domain variation of HAP, we obtain a hydrophobic stress tensor which is given as follows,

\begin{equation}
\label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u\otimes \nabla u\right).
\end{equation}
%
The integration over boundaries gives force and torque due to the
hydrophobic attraction 
\begin{align}
  \label{eqn:hydrophobicAttraction}
  \FF_i = \int_{\bd \Gamma_i} {\bf T}\cdot \nnu_i \, dS, 
    \quad  \text{ and } \quad
  T_i = \int_{\Gamma_i} (\xx - \aa_i) \times ({\bf T} \cdot \nnu_i) dS,
\end{align}
%
where $\xx \in \Gamma_i$, $\aa_i$ is the $i^{th}$ particle center, and
$\nnu_i$ is the outward normal of the $i^{th}$ particle.

%\begin{equation}
%        \label{SL}
%\begin{cases}
% -\rho^2 \Delta u + u =0 & \text{ in } \Omega,\\
%  u(x) = f(x)  &\text{ on }\Sigma, \\
%   u(x) \to 0, &\text{ as } x \to \infty.
%\end{cases}
%\end{equation}

\subsection{Nondimensionalizations}

@@

\subsection{\label{mobility}Mobility Problem}

The governing equations of $N_b$-body bodies suspended in the unbounded
domain $\Omega$ are
\begin{alignat}{3}
  -\Delta \uu + \mu\nabla p, &= 0 \qquad \qquad && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu &\to \uu_\infty, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $p$ is the pressure, and $\uu_\infty$ is
the background flow. Each particle $\Gamma_i$ with center $\aa_i$ is
assumed to be a rigid body, so its two-dimensional velocity can be
written as
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity. Therefore, the no-slip boundary condition of
the velocity field is 
\begin{align}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \bd\Gamma_i.
\end{align}
To determine the translational and angular velocities of each rigid
body, we impose a force balance and torque balance. Matching the total
force and torque in~\eqref{eqn:hydrophobicAttraction} with the
hydrodynamic force, we require
\begin{align}
  \label{eqn:force}
  \FF_i &= \int_{\bd\Gamma_i} \ssigma \cdot \nnu_i \, ds, \\
  \label{eqn:torque}
  T_i &= \int_{\bd\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu_i) \, ds,
\end{align}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor.

\subsection{Short Range Repulsion}

At each time step, we examine the nearest grid points between each pair of body  boundaries. Within a specified distance $d$, we introduce a short range repulsive potential as follows

\begin{equation}
\Phi_{repul} = 1- \sin\left(\frac{\pi}{2} d\right).
\end{equation}

Here we would like to make a remark about the time marching scheme. From the mobility problem, we obtain velocity and torque profiles of an $N_b$-body system. A second order Adam-Bashforth scheme is used for updating the dynamics but body collisions may occur during the simulation while a relatively large time step is chosen. We therefore impose a short range repulsion that it will be turned on when the distance between two boundary points on different bodies are within one characteristic length.

@@

\begin{figure}
\centering
\includegraphics[height=1.2in]{N116_shear_0.pdf}
\includegraphics[height=1.2in]{N116_shear_3000.pdf}
\includegraphics[height=1.2in]{N116_shear_6000.pdf}
\includegraphics[height=1.2in]{N116_shear_9000.pdf}\\
  \caption{Two vesicles under a shear flow where the dimensioneless shear rate is $\dot\gamma=0.005$; The initial configuration is shown in panel (a). The background color indicates the magnitude of fluid pressure. From dark red to dark blue, it represents the the pressure varies from high to low values. The color bar is valid for all panels. Streamlines are plotted in purple.
  }
    \label{figure9}
\end{figure}

>>>>

% Bryan's Marcros
\renewcommand{\aa}{\mathbf{a}}
\newcommand{\bd}{\partial}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\eeta}{\boldsymbol{\eta}}
\newcommand{\FF}{\mathbf{F}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\nnu}{\boldsymbol{\nu}}
\newcommand{\ssigma}{\boldsymbol{\sigma}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\RR}{\mathbf{R}}
\renewcommand{\SS}{\mathbf{S}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\yy}{\mathbf{y}}

@@

\author{
Szu-Pei Fu\aff{1},
Bryan Quaife\aff{2},
Rolf Ryham\aff{1}, \and
Yuan-Nan Young\aff{3}
}
 \affiliation{
\aff{1}Department of Mathematics, \\Fordham University, Bronx, New York 10458, USA
\aff{2}Department of Scientific Computing, \\Florida State University, Tallahassee, Florida 32306, USA
\aff{3}Department of Mathematical Sciences, New Jersey Institute of Technology,\\ Newark, New Jersey 07102, USA
 }

@@

\section{Governing  Equations}
\subsection{\label{mobility}Mobility Problem}
The governing equations of $N_b$-body particles suspended in the
two-dimensional, unbounded
domain $\Omega$ are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= 0 \qquad \qquad && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $\mu$ is viscosity, $p$ is the pressure, and $\uu_\infty$ is
the background flow. Each particle $\Gamma_i$ with center $\aa_i$ is
a rigid body, and its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity.
Here, $\langle x, y \rangle^{\perp} = \langle -y, x \rangle$.
Therefore, the no-slip boundary condition of
the velocity field is 
\begin{align}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \bd\Gamma_i.
\end{align}
To determine the translational and angular velocities of each rigid body,
we assume some imposed forces $\FF_i$ and torques $T_i$ acting on each particle.
Since the small particles are intertialess, force balance gives 
\begin{alignat}{2}
  \label{eqn:force}
  \FF_i &+ \int_{\bd\Gamma_i} \ssigma \cdot \nnu_i \, ds && = \mathbf{0}, \\
  \label{eqn:torque}
  T_i &+ \int_{\bd\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu_i) \, ds && = 0,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor).

\subsection{Imposed Forces}
The imposed forces and torques contain two parts. 
The first part comes from the hydrophobic attraction potential (HAP).
This potential was introduced in \cite{Fu20} and
models hydrophobic attraction by solving the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{SL}
-\rho^2 \Delta u + u &=0,            && \mathbf{x} \in \Omega,\\
\label{SLbc}
u(\mathbf{x}) &= f_i(\mathbf{x}),\qquad  && \mathbf{x} \in \Gamma_i, \\
\label{SLff}
u &\to 0,                          &&|\mathbf{x}| \to \infty,
\end{alignat}
where $f_i$ is material label with $f_i = 0$, resp. $1$ representing hydrophilic, resp. hydrophobic, 
portions of the surface, and $\rho$ is the decay length of attraction.
The forces and torques for this attraction are 
\begin{align}
  \label{eqn:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\bd \Gamma_i} {\bf T}\cdot \nnu_i \, dS, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu_i) dS,
\end{align}
where
\begin{equation}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u\otimes \nabla u\right)
\end{equation}
is the hydrophobic stress tensor.  The parameter $\gamma$ is interfacial tension. 
The forces $\FF_i^{\text{hydro}}$ and torques $T_i^{\text{hydro}}$ are responsible for
forming particle aggregates that partially sequester 
the hydrophobic portions of the particle surface.

The second part of the imposed forces and torques comes from repulsion between proximal particles.
We define 
\begin{equation}
  \label{REPUL}
  \FF_i^{\text{repul}} = \sum_{i=1}^{N_b} \frac{\mathbf{x}_{ij} - \mathbf{y}_{ij}}{\rho_{ij}} P(\rho_{ij}),\quad 
  T_i^{\text{repul}} = \sum_{i=1}^{N_b} (\mathbf{x}_{ij} - \mathbf{a}_i)^{\perp}\cdot \frac{\mathbf{x}_{ij} - \mathbf{y}_{ij}}{\rho_{ij}} P(\rho_{ij})
\end{equation}
where $\mathbf{x}_{ij} \in \partial \Gamma_i$, 
$\mathbf{y}_{ij} \in \partial \Gamma_j$, 
and $\rho_{ij} = |\mathbf{x}_{ij} - \mathbf{y}_{ij}|$
is the distance between the pair of particles.
The repulsion profile $P(r)$ is zero for distances $r$ larger than a repulsion length scale
$\rho_0$.  As such, \eqref{REPUL} ignores particles $\Gamma_j$ outside a tubular neighborhood
of $\Gamma_i$.  For $0 \leq r < \rho_0,$ we use $P(r) = M(1 - \sin(r/\rho_0))$ where $M$ is sufficiently
large to prevent particle collisions.  The total imposed force and torque are
\begin{equation}
  \FF_i = \FF_i^{\text{hydro}} + \FF_i^{\text{repul}},\quad
    T_i =   T_i^{\text{hydro}} +   T_i^{\text{repul}}.
\end{equation}

\subsection{Time marchning and Nondimensionalizations}
From the mobility problem, we obtain translational and angular velocities of the $N_b$-body system.
A second order Adam-Bashforth scheme updates the particle positions and orientations.
dynamics but body collisions may occur during the simulation while a relatively large time step is chosen.
We therefore impose a short range repulsion that it will be turned on when the distance between two boundary points on different bodies are within one characteristic length.

@@

@@

\begin{figure}
\centering
\includegraphics[height=1.2in]{N116_shear_0.pdf}
\includegraphics[height=1.2in]{N116_shear_3000.pdf}
\includegraphics[height=1.2in]{N116_shear_6000.pdf}
\includegraphics[height=1.2in]{N116_shear_9000.pdf}\\
  \caption{Two vesicles under a shear flow where the dimensioneless shear rate is $\dot\gamma=0.005$; The initial configuration is shown in panel (a). The background color indicates the magnitude of fluid pressure. From dark red to dark blue, it represents the the pressure varies from high to low values. The color bar is valid for all panels. Streamlines are plotted in purple. The insets of all panels are generated from the results of continuum modeling simulations and the streamlines have a perfect agreement to the numerical results.
  }
    \label{figure9}
\end{figure}
